---
title: "Data Description"
author: "Le Thi Thanh Tam _ 202005018"
date: "1/16/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Prescription (OD_RX file _Treatment data)

## Open and rename OD_RX file
```{r df}
load("E:/Practicum 2/OD_RX.RData")
RXdf = df
```
## Description of RXdf data
* Description of the market is "Oral Diabetes"
```{r}
unique(RXdf$MARKET_DESC)
```

* The data contain 182,680 patient visits where a prescription was written by 2,127 physicians
```{r}
# Numbers of patient visits
length(unique(RXdf$PATIENT_VISIT_ID))
# Numbers of Physicians who prescrip products
length(unique(RXdf$GEO_PHYSICIAN_KEY))
```

* The data of treatment specializing in the pharmaceutical industry. This data was conducted for over an 18 month period from January 1st, 2007 to June 30th, 2008
```{r}
# Check the date period
summary(RXdf$DAY_DT)
```

* During this time span, physicians report the name of the product prescribed during each patient visit
```{r}
# Showing the name of prescribed product
unique(RXdf$PRODUCT)
length(unique(RXdf$PRODUCT))
```

* There are two types of description of diagnosis (new diagnosis or ongoing diagnosis). For prescription switches brand, there has a short written description of the reason for switching is provided. In addition, the data includes the patient requested for a particular product

* Top 5 products were prescribed most frequently are *Glucophage/Metformin*, *Actos*, *Sulfonylureas*, *Other Insulins*, *Lantus*
```{r}
Pro = table(RXdf$PRODUCT)
head(sort(Pro, decreasing = TRUE), n=5)
```

# Detailing (OD_Detailing file_Detail Data)
## Open and rename OD_Detailing data

```{r}
load("E:/Practicum 2/OD_Detailing.RData")
DTdf = df
```

## Description of DTdf data
* The data contains 1,958 Physician number for 56,685 detailing visits by pharmaceutical salespeople
```{r}
# Physician number
length(unique(DTdf$GEO_PHYSICIAN_KEY))
# Sales visit to identify unique visit
length(unique(DTdf$SALES_VISIT_ID))
```

# A table for Prescription and Detailing for Products

```{r}
table1 = table(RXdf$PRODUCT)
table2 = table(DTdf$PRODUCT)
p1 = prop.table(table1)
p2 = prop.table(table2)
table = data.frame(table1 = table1, p1=p1, table2=table2, p2=p2)
colnames(table) = c("Product", "Prescription", "Product2", "Percentage(%)",
                    "Product3", "Detailing", "Product4", "Percentage(%)")
table$Product2 = NULL
table$Product3 = NULL
table$Product4 = NULL
knitr::kable(table,caption = "Prescription and Detailing Shares for the Drugs", 
             "pipe", digits = 3,
             align = c("l", "c", "c","c","c"))
```

* From the table, the top prescription share products are *Glucophage/Metformin*, *Actos*, *Sulfonylureas*, *Other Insulins*, *Lantus*. Glucophage/Metformin is the largest prescription share drug (0.26%), however it accounts for the smallest percentage in detailing. Other Insulins accounts for the highest detailed product in this data. 
